Information for DOE Report

### 2015-11-29 DOESummaryNov2015/DOESummaryNov2015.tex
First finished writeup.

\subsection{Comparison of Photon Detector Designs at CDDF}
We compared the performance of the photon detector designs from CSU and LSU. Detectors were put side by side in the Dewar at CDDF. Each detector has 2 SiPM channels. Figure~\ref{fig:cddfadc} shows the integrated ADC for the four SiPM channels. A threshold scan was done to set the threshold just above the pedestal peak.

@@

Figure~\ref{fig:cddfcalibrated} shows calibrated photoelectron(PE) spectra for the four channels. The calibration was done by assigning the first peak to 1 PE, the second to 2 PE, etc.

@@

>>>>

\subsection{Comparison of Photon Detector Designs at CDDF}
We compared the performance of the photon detector designs from CSU and LSU. Detectors were put side by side in the Dewar at CDDF. Each detector has 2 SiPM channels. Figure~\ref{fig:cddfadc} shows the integrated ADC for the four SiPM channels. A threshold scan was done to set the threshold just above the pedestal peak. The detectors were irradiated by an $\alpha$ source and triggered at 0.5 photoelectron for each channel independently.

@@

\begin{figure}
  \centering
  \includegraphics[angle=270,origin=c,width=.7\textwidth]{figures/avgPE.pdf}
  \caption{Average light yield for at least 2 PEs for the four channels.}
  \label{fig:cddfly}
\end{figure}

Figure~\ref{fig:cddfcalibrated} shows calibrated photoelectron (PE) spectra for the four channels. The calibration was done by assigning the first peak to 1 PE, the second to 2 PE, etc.

Since the data showed a high trigger rate, the light yields were compared by excluding the single PE events. Figure~\ref{fig:cddfly} shows the average number of photoelectrons of events with at least 2 PEs. This figure shows the four channels have comparable light yield in this $\ge 2$ PE measure.

\subsection{Monitor Application for Voltage Readback of the SiPM Signal Processor}

CSU owns 2 SiPM Signal Processors (SSP). The SSP supplies high voltage to its 12 independent channels, and for each channel a readback register is in place for monitoring voltage in operation. A monitor application is developed to serve this purpose.

\subsection{Control Application for the LBNE Calibration Module}

This module is named after the former name of the experiment, LBNE. The LBNE Calibration Module (LCM) outputs calibration signals to 3 different calibration subsystem, namely, the Indiana University (IU) system, the TPC system, and the photon detector (PD) system. Figure~\ref{fig:pulser_pulse} shows sample pulses generated by the control application for the LCM. To facilitate the use of the control application, a simple GUI frontend is also provided as shown in Figure~\ref{fig:pulser_frontend}.

\begin{figure}
  \centering
  \includegraphics[width=.7\textwidth]{figures/pulser.jpg}
  \caption{Sample pulses generated by the control application for the LCM.}
  \label{fig:pulser_pulse}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.7\textwidth]{figures/Screenshot_from_2015-09-23_10-49-58.png}
  \caption{A GUI frontend for the control application for the LCM.}
  \label{fig:pulser_frontend}
\end{figure}

@@

\subsection{Generation of Flat Monte Carlo Neutrino Spectrum}

The neutrino flux of the NuMI beam peaks at 2 GeV. The total interaction cross-section of neutrino is roughly $\sim E$, which at high energy is dominated by the deep inelastic scattering. The left plot in Figure~\ref{fig:NuMISpec} shows the total neutrino flux of the NuMI beam (the black histogram) from simulation. The right plot in Figure~\ref{fig:NuMISpec} is the neutrino event spectrum. The event spectrum is highly peaked around 2 GeV, leading to lack of statistics for particle identification (PID) training with neural network algorithms outside the peak range. There is need to warp the input neutrino flux to have a flat neutrino event spectrum for the PID training of the off-peak events. The idea is to make the input flux $\sim1/E$. Figure~\ref{fig:inverseEflux} shows the $1/E$ warped flux. Figure~\ref{fig:flatspec} is a sample output of the neutrino event spectrum gone through GENIE simulation with a $1/E$ input flux, which has an evenly distributed number of events in the interested energy range.

\begin{figure}
  \centering
  \includegraphics[width=.7\textwidth]{figures/NuMI_spectrum.jpg}
  \caption{Left: The total neutrino flux of NuMI beam (black) and the fluxes of various components. Right: The neutrino event spectra for the corresponding fluxes.}
  \label{fig:NuMISpec}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.4\textwidth]{figures/inverseEflux.png}
  \caption{The warped $1/E$ flux.}
  \label{fig:inverseEflux}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.7\textwidth]{figures/warped_event_spectrum.eps}
  \caption{The neutrino event spectrum with a $1/E$ input flux.}
  \label{fig:flatspec}
\end{figure}

\subsection[Numu on e Event Topology]{$\nu_\mu$ on e event topology}

$\nu_\mu-e$ scattering is one of the simplest processes whose cross-section can be calculated very accurately. Therefore these events can be used for constraining the NuMI flux. The characteristic feature of the $\nu_\mu-e$ events leaving in the detector is a 1-prong electromagnetic shower in the very forward direction.  We used the near detector MC data to check how well our current reconstruction algorithms work for $\nu_\mu-e$ scattering events. A sample event where the current reconstruction algorithms reconstruct successfully as a 1-prong event is given in Figure~\ref{fig:oneprong}. However there are also events in which the number of the reconstructed prongs is more than one, such as that shown in Figure~\ref{fig:multiprong}. This is because current algorithms pull the vertex downstream into the shower, leaving 2 back-to-back prongs. The proportion of events for which current algorithms perform well is about $70\%$, while the proportion of the incorrectly reconstructed events is about $30\%$, leaving room for improvements.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/nelastic1npng3d1.eps}
  \caption{A $\nu_\mu$ on $e$ event where the current reconstruction algorithms reconstruct successfully as a 1-prong event. The upper plot is the top view of the NO$\nu$A near detector, while the bottom is the side view of it.}
  \label{fig:oneprong}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/nelastic1npng3d1+.eps}
  \caption{A $\nu_\mu$ on $e$ event where the current reconstruction algorithms reconstruct incorrectly as a multi-prong event. The hollow cross indicates the reconstructed vertex where the interaction happens. Current algorithms pull the vertex downstream into the shower, leaving 2 back-to-back prongs.}
  \label{fig:multiprong}
\end{figure}
